---
title: "2. Software" # A Brief Introduction to R and SAS"  
output: 
  html_document:
    toc: true
    fig_caption: true
bibliography: bibliography.bib    
---

### Reading 

* @shumway/stoffer:2011
  
  
# Brief Introduction to R

I recommend you download and install R from [CRAN](https://cran.r-project.org/) and [R Studio](https://www.rstudio.com/).

Have a look at a number of online tuturials such as [this](https://www.otexts.org/fpp/using-r) and familiarize yourself with the R syntax.

Displaying versus not displaying results

```{r}
x <- 3
x = 3
y = 5
(y = 5)

```

Here is some simple things you can do in R

```{r}
# addition

1 + 2

# multiplication and addition

2*3 - 10

# division and subtraction

4/4 - 3

# log, exponential, pi

log(exp(pi))

# sinusoids

sin(pi/2)

# power

exp(1)^(-2)

# square root

sqrt(8)

# sequences

1:10

seq(1,20,by=2)

rep(3,9)

```

```{r}
myData = c(1,2,3,2,1)
myTimeSeries = as.ts(myData)
```


  Right     Left     Center     Default
-------     ------ ----------   -------
     12     12        12            12
    123     123       123          123
      1     1          1             1

A lot of this course will be to learn some `code`.

__Some Matrices__

```{r}
library(MTS)
library(mvtnorm)
diag(10)
sig = diag(2)
x = rmvnorm(300,rep(0,2),sig)
```

```{r}
data("mts-examples", package="MTS")
```

We want to load the library "TSA" from @cryer/chan:2008
```{r}
# install.packages("TSA")
library(TSA)
```

# A. APPS.SMU

Apps.smu.edu and accessing computer programs on SMU's virtual server - downloading Citrix Receiver

# B. Introduction to SAS (SAS = Statistical Analysis System)

## i. Program Editor in SAS 9.4

## ii. Data Steps and Procedure Steps

## iii. Log and Listing Files

# C. Inputting Data

```{r, echo=FALSE}
library(xts)
library(forecast)
library(quantmod)
format(as.yearqtr(Sys.Date()), "%Y Q%q")
```

```{r}
# downloading the FRED data directly, transforming to quarterly, from 1985

un <- to.quarterly(getSymbols("UNRATE",src="FRED",auto.assign=FALSE,return.class="xts")['1985/'], OHLC = FALSE)  
rgdp_aqoq <- to.quarterly(getSymbols("A191RL1Q225SBEA",src="FRED",auto.assign=FALSE,return.class="xts")['1985/'], OHLC = FALSE) 
pce_level <- to.quarterly(getSymbols("PCECTPI",src="FRED",auto.assign=FALSE,return.class="xts")['1985/'], OHLC = FALSE)
pce_core_yoy <- to.quarterly(getSymbols("BPCCRO1Q156NBEA",src="FRED",auto.assign=FALSE,return.class="xts")['1985/'], OHLC = FALSE)
ffr <- to.quarterly(getSymbols("FEDFUNDS",src="FRED",auto.assign=FALSE,return.class="xts")['1985/'], OHLC = FALSE)

macro.data <- merge.xts(pce_core_yoy,pce_level,rgdp_aqoq,un,ffr)

names(macro.data) <- c("p_core","p_head","dy","un","ffr")
names(macro.data)
head(macro.data)
tail(macro.data)

```
  
## i. Direct Input

## ii. Infile Statement

* Textbooks: 
  + @cowpertwait/metcalfe:2009
  + @luetkepohl:2005

# References