---
title: "3. Salient Characteristics of Time Series Data"  
output: 
  html_document:
    toc: true
    fig_caption: true
bibliography: bibliography.bib      
---
  
# A. Trend, Seasonal, Cycle, Irregular Components

Let us first look at an example of a time-series. 

```{r, eval=FALSE}
ts <- AirPassengers
plot(AirPassengers)

```

```{r}
plot(AirPassengers,las=1,col="blue",main="Air Passenger Data")

```

Now let us consider seasonal plots

```{r}
seasonplot(AirPassengers,las=1,main="Seasonal Plot")
```


```{r}
monthplot(AirPassengers,las=1,main="Month Plot")
```


Another example is the CO2 data see `?monthplot`

```{r}
fit <- stl(log(co2), s.window = 20, t.window = 20)
plot(fit)
op <- par(mfrow = c(2,2))
monthplot(co2, ylab = "data", cex.axis = 0.8)
monthplot(fit, choice = "seasonal", cex.axis = 0.8)
monthplot(fit, choice = "trend", cex.axis = 0.8)
monthplot(fit, choice = "remainder", type = "h", cex.axis = 0.8)
par(op)
```


$$ \{ y(s,t), s \in \mathbb{S}, t \in \mathbb{T}\}$$ 
What is a stationary stochastic process?


$$\begin{align*} 
y'_t &= \beta_1 x'_{1,t} + \dots +
\beta_k x'_{k,t} + n'_t,\\ 
& (1-\phi_1B)n'_t = (1+\theta_1B)e_t, \end{align*} $$

where $y'_t=y_t-y_{t-1}$, $x'_{t,i}=x_{t,i}-x_{t-1,i}$ and $n'_t=n_t-n_{t-1}$, which is
a regression model in differences with ARMA errors.


# B. Y = T + S + C + I (Additive Decomposition)

$$ Y = T + S + C + I $$

# C. log(Y) = T + S + C + I (Multiplicative Decomposition)

# D. A Stylized Decomposition of a Time Series

# E. Importance of knowning which components are in your time series

# References